### 2021-08-05 ch02_scientific_background/scientific_background.tex
Small linguistic optimizations

A possible way out is given by numerical climate models, which offer the
possibility to simulate the Earth's climate on a computer. To efficiently
replicate the vastly complex Earth system with finite computational resources,
climate models divide the Earth into a set of grid cells. The typical size of
an atmospheric grid cell in modern-day global climate models is about $100
\unit{km}$ along the horizontal dimensions (latitude and longitude) and $1
\unit{km}$ along the vertical dimension (pressure level or height). In
addition, also the temporal evolution of the Earth system is discretized using
time steps that are typically about $30 \unit{min}$ long in modern-day global
climate models. For each grid cell and time step, a single value per model
variable is usually given. Examples for such variables provided by a climate
model are the atmospheric prognostic variables velocity (horizontal and
vertical), temperature, specific humidity, pressure, and density. These
prognostic variables are related to each other via the \emph{primitive
  equations}, a set of partial differential equations that can be derived from
the conservation of momentum, mass, energy, and moisture \autocite{Holton2004}.
To progress further in time, these primitive equations are solved numerically
by the dynamical core of the climate models, which eventually describes the
Earth's large-scale atmospheric motions. Similar to this, many other processes
of the Earth system are simulated by the climate models using other fundamental
physical laws and principles. As opposed to these calculated variables that
form the output of climate models, the corresponding input is mainly given by
external forcings, which are incorporated as boundary conditions into the
climate models. For example, the major driver of anthropogenic climate change,
the \ac{GHG} \ac{CO2}, is either included with prescribed atmospheric \ac{CO2}
concentrations (\emph{concentration-driven runs}) or prescribed \ac{CO2}
emissions (\emph{emission-driven runs}).

Many processes in the Earth system occur at spatial scales much smaller than
the size of a typical grid cell. Illustrative examples for this are clouds,
which are usually smaller than a $100 \unit{km} \times 100 \unit{km}$ grid
cell, but still play an important role in the overall climate system by
reflecting incoming and outgoing radiation \autocite{Boucher2013}. To
reasonably approximate these subgrid-scale processes, a concept called
\emph{parameterization} is used. Instead of simulating a process exactly,
parameterizations aim to represent the effect of that process at the grid scale
of the climate model by generating the appropriate forcing terms for the rest
of the system and the rest of the processes \autocite{Gettelman2016}.
Parameterizations are necessary to simulate many processes of the Earth system,
for example, surface heat and moisture fluxes, moist convection, turbulent
mixing, and radiation \autocite{Holton2004}.

@@

The second source of uncertainty in climate model projections is \emph{emission
  uncertainty}. This arises from the different possible trajectories in terms
of future forcing (\acp{GHG}, aerosols, land-use changes, \etc{}) humankind
might take. Examples for these are the \acs{SSP}-based experiments given by
\ac{ScenarioMIP} that include a variety of different scenarios from a
sustainable future to a full fossil fuel--based development (see
\cref{subsec:02:cmip}). A possible approach to quantify emission uncertainty is
to assess the climate impact of these different trajectories. Since the
emission uncertainty strongly depends on the future development of human
society, it cannot be reduced by improving climate models. In contrast to
natural variability, the emission uncertainty increases over time in climate
projections, since estimating forcings for the near future is simpler than for
the far future.

@@

For some domains of the Earth system, feedbacks can be positive and/or
negative, since a variety of different mechanisms are involved. An example of
this is the cloud feedback. Changes in the clouds induced by climate change can
cause changes in their longwave (greenhouse warming) and shortwave (reflective
cooling) effects on the Earth's radiation budget, which both need to be
considered for the overall cloud feedback \autocite{Boucher2013}. Relevant
cloud properties that may change as a response to an external forcing and that
may alter the Earth's radiative budget are cloud cover, cloud optical
thickness, cloud top and cloud base height, vertical extent, and the
geographical distribution of clouds. Examples for possible cloud feedback
processes are the increase in cloud top height of high-level clouds in a
warming climate which traps longwave radiation and enhances global warming, and
the reduction in mid- and low-level cloud cover which diminishes the reflection
of incoming solar radiation and also increases the surface warming
\autocite{Boucher2013}. In global climate model ensembles, the overall cloud
feedback shows a large range with positive and negative values, but tends to be
slightly positive on average \autocite{Soden2006, Dufresne2008, Vial2013,
  Zelinka2020}. This large uncertainty in the cloud feedback is a major reason
for uncertainties in the climate sensitivity of climate models
\autocite{Boucher2013, Flato2013}.

@@

The different land carbon cycle models integrated into modern \acp{ESM} use a
variety of different parameterizations for the different processes
\autocite{Arora2020}. Examples for photosynthesis (\ac{GPP}) are given by
\textcite{Farquhar1980} for \ch{C3} plants and \textcite{Collatz1992} for
\ch{C4} plants. The terms \enquote{\ch{C3}} and \enquote{\ch{C4}} refer to the
different metabolic pathways used in carbon fixation during photosynthesis. In
both cases, the enzyme \ac{RuBisCO} is an important catalyst for the
corresponding chemical reactions. Both parameterizations utilize variables that
are known to affect photosynthesis, such as the availability of light, \ac{CO2}
concentration, soil moisture, and temperature.

>>>>

A possible way out is given by numerical climate models, which offer the
possibility to simulate the Earth's climate on a computer. To efficiently
replicate the vastly complex Earth system with finite computational resources,
climate models divide the Earth into a set of grid cells. The typical size of
an atmospheric grid cell in modern-day global climate models is about $100
\unit{km}$ along the horizontal dimensions (latitude and longitude) and $1
\unit{km}$ along the vertical dimension (pressure level or height). In
addition, also the temporal evolution of the Earth system is discretized using
time steps that are typically about $30 \unit{min}$ long in modern-day global
climate models. For each grid cell and time step, a single value per model
variable is usually given. Examples of such variables provided by a climate
model are the atmospheric prognostic variables velocity (horizontal and
vertical), temperature, specific humidity, pressure, and density. These
prognostic variables are related to each other via the \emph{primitive
  equations}, a set of partial differential equations that can be derived from
the conservation of momentum, mass, energy, and moisture \autocite{Holton2004}.
To progress further in time, these primitive equations are solved numerically
by the dynamical core of the climate models, which eventually describes the
Earth's large-scale atmospheric motions. Similar to this, many other processes
of the Earth system are simulated by the climate models using other fundamental
physical laws and principles. As opposed to these calculated variables that
form the output of climate models, the corresponding input is mainly given by
external forcings, which are incorporated as boundary conditions into the
climate models. For example, the major driver of anthropogenic climate change,
the \ac{GHG} \ac{CO2}, is either included with prescribed atmospheric \ac{CO2}
concentrations (\emph{concentration-driven runs}) or prescribed \ac{CO2}
emissions (\emph{emission-driven runs}).

Many processes in the Earth system occur at spatial scales much smaller than
the size of a typical grid cell. Illustrative examples of this are clouds,
which are usually smaller than a $100 \unit{km} \times 100 \unit{km}$ grid
cell, but still play an important role in the overall climate system by
reflecting incoming and outgoing radiation \autocite{Boucher2013}. To
reasonably approximate these subgrid-scale processes, a concept called
\emph{parameterization} is used. Instead of simulating a process exactly,
parameterizations aim to represent the effect of that process at the grid scale
of the climate model by generating the appropriate forcing terms for the rest
of the system and the rest of the processes \autocite{Gettelman2016}.
Parameterizations are necessary to simulate many processes of the Earth system,
for example, surface heat and moisture fluxes, moist convection, turbulent
mixing, and radiation \autocite{Holton2004}.

@@

The second source of uncertainty in climate model projections is \emph{emission
  uncertainty}. This arises from the different possible trajectories in terms
of future forcing (\acp{GHG}, aerosols, land-use changes, \etc{}) humankind
might take. Examples of these are the \acs{SSP}-based experiments given by
\ac{ScenarioMIP} that include a variety of different scenarios from a
sustainable future to a full fossil fuel--based development (see
\cref{subsec:02:cmip}). A possible approach to quantify emission uncertainty is
to assess the climate impact of these different trajectories. Since the
emission uncertainty strongly depends on the future development of human
society, it cannot be reduced by improving climate models. In contrast to
natural variability, the emission uncertainty increases over time in climate
projections, since estimating forcings for the near future is simpler than for
the far future.

@@

For some domains of the Earth system, feedbacks can be positive and/or
negative, since a variety of different mechanisms are involved. An example of
this is the cloud feedback. Changes in the clouds induced by climate change can
cause changes in their longwave (greenhouse warming) and shortwave (reflective
cooling) effects on the Earth's radiation budget, which both need to be
considered for the overall cloud feedback \autocite{Boucher2013}. Relevant
cloud properties that may change as a response to an external forcing and that
may alter the Earth's radiative budget are cloud cover, cloud optical
thickness, cloud top and cloud base height, vertical extent, and the
geographical distribution of clouds. Examples of possible cloud feedback
processes are the increase in cloud top height of high-level clouds in a
warming climate which traps longwave radiation and enhances global warming, and
the reduction in mid- and low-level cloud cover which diminishes the reflection
of incoming solar radiation and also increases surface warming
\autocite{Boucher2013}. In global climate model ensembles, the overall cloud
feedback shows a large range with positive and negative values, but tends to be
slightly positive on average \autocite{Soden2006, Dufresne2008, Vial2013,
  Zelinka2020}. This large uncertainty in the cloud feedback is a major reason
for uncertainties in the climate sensitivity of climate models
\autocite{Boucher2013, Flato2013}.

@@

The different land carbon cycle models integrated into modern \acp{ESM} use a
variety of different parameterizations for the different processes
\autocite{Arora2020}. Examples of photosynthesis (\ac{GPP}) are given by
\textcite{Farquhar1980} for \ch{C3} plants and \textcite{Collatz1992} for
\ch{C4} plants. The terms \enquote{\ch{C3}} and \enquote{\ch{C4}} refer to the
different metabolic pathways used in carbon fixation during photosynthesis. In
both cases, the enzyme \ac{RuBisCO} is an important catalyst for the
corresponding chemical reactions. Both parameterizations utilize variables that
are known to affect photosynthesis, such as the availability of light, \ac{CO2}
concentration, soil moisture, and temperature.

### 2021-08-05 ch03_esmvaltool/esmvaltool.tex
Small linguistic optimizations

Since its first release in 2016 \autocite{Eyring2016a}, the \ac{ESMValTool} has
been greatly extended and improved. A schematic representation of the current
structure of the \ac{ESMValTool} is illustrated in
\cref{fig:03:esmvaltool_schematic}. The first major element of this workflow is
an extensive preprocessing of the input data. For this, the Python-based
\acf{ESMValCore} provide a set of operators (\emph{preprocessors}) that are
applied to the input data. There are two classes of preprocessors: non-optional
and optional preprocessors. Examples for non-optional preprocessors are the
input/output operations \emph{load} (loading input data from climate models and
observations) and \emph{save} (saving the preprocessed data), which utilize the
\ac{NetCDF} format (a self-describing, machine-independent binary file format
used for the storage of array-based scientific data). Further non-optional
preprocessors include \emph{checks} and \emph{fixes} of the input data, which
test whether the input data adheres to standards given by the \ac{CMOR} format
and fixes the data if necessary. The \ac{CMOR} format ensures that the output
from the many different modeling institutes within a generation of \ac{CMIP}
follows common standards. The optional preprocessors include commonly used
operations on spatiotemporal datasets, like spatial and temporal
\emph{subsetting}, the calculation of \emph{spatial and temporal statistics},
horizontal and vertical \emph{interpolation}, land/sea/ice \emph{masking},
\emph{unit conversion}, and the calculation of \emph{multi-model statistics}. A
further example is \emph{variable derivation}, which can be used to derive
non-\ac{CMOR} variables from the input data. The preprocessing functions of
\ac{ESMValCore} aim to facilitate the routine evaluation of climate models by
providing a set of commonly used data operations. To reduce computation times,
\ac{ESMValCore} allows parallel preprocessing of different datasets. However,
since this leads to higher memory usage, a compromise has to be found in
practice to optimize the use of time and memory resources.

>>>>

Since its first release in 2016 \autocite{Eyring2016a}, the \ac{ESMValTool} has
been greatly extended and improved. A schematic representation of the current
structure of the \ac{ESMValTool} is illustrated in
\cref{fig:03:esmvaltool_schematic}. The first major element of this workflow is
an extensive preprocessing of the input data. For this, the Python-based
\acf{ESMValCore} provide a set of operators (\emph{preprocessors}) that are
applied to the input data. There are two classes of preprocessors: non-optional
and optional preprocessors. Examples of non-optional preprocessors are the
input/output operations \emph{load} (loading input data from climate models and
observations) and \emph{save} (saving the preprocessed data), which utilize the
\ac{NetCDF} format (a self-describing, machine-independent binary file format
used for the storage of array-based scientific data). Further non-optional
preprocessors include \emph{checks} and \emph{fixes} of the input data, which
test whether the input data adheres to standards given by the \ac{CMOR} format
and fixes the data if necessary. The \ac{CMOR} format ensures that the output
from the many different modeling institutes within a generation of \ac{CMIP}
follows common standards. The optional preprocessors include commonly used
operations on spatiotemporal datasets, like spatial and temporal
\emph{subsetting}, the calculation of \emph{spatial and temporal statistics},
horizontal and vertical \emph{interpolation}, land/sea/ice \emph{masking},
\emph{unit conversion}, and the calculation of \emph{multi-model statistics}. A
further example is \emph{variable derivation}, which can be used to derive
non-\ac{CMOR} variables from the input data. The preprocessing functions of
\ac{ESMValCore} aim to facilitate the routine evaluation of climate models by
providing a set of commonly used data operations. To reduce computation times,
\ac{ESMValCore} allows parallel preprocessing of different datasets. However,
since this leads to higher memory usage, a compromise has to be found in
practice to optimize the use of time and memory resources.
